### 2019-04-29 report/MSc_Report.tex
Presentation upload

@@

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Thu Apr 25 14:26:47 2019
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{lllrrrr}
  \hline
Study & Comparison & Outcome & Events & Total & Events\_c & Total\_c \\ 
  \hline
Bohn 1989 & Barbiturate vs no barbiturate & Death at the end of follow-up & 11 & 41 & 11 & 41 \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Death at the end of follow-up & 14 & 27 & 13 & 26 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Example of two results as given in the dataset. Events denotes the count of events in the treatment group while Events c the count of events in the group compared to. Further descriptive variables have been ommitted} 
\label{barbiturate.row}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Thu Apr 25 14:26:47 2019
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lll}
  \hline
Study & Comparison & Outcome \\ 
  \hline
Bohn 1989 & Barbiturate vs no barbiturate & Death at the end of follow-up \\ 
  Bohn 1989 & Barbiturate vs no barbiturate & Death or severe disability at the end of follow-up \\ 
  Eisenberg 1988 & Barbiturate vs no barbiturate & Death at the end of follow-up \\ 
  Eisenberg 1988 & Barbiturate vs no barbiturate & Uncontrolled ICP during treatment \\ 
  Eisenberg 1988 & Barbiturate vs no barbiturate & Hypotension during treatment \\ 
  Perez-Barcena 2008 & Pentobarbital vs Thiopental & Death at the end of follow-up (6 months) \\ 
  Perez-Barcena 2008 & Pentobarbital vs Thiopental & Death or severe disability at the end of follow-up (6 months) \\ 
  Perez-Barcena 2008 & Pentobarbital vs Thiopental & Uncontrolled ICP during treatment \\ 
  Perez-Barcena 2008 & Pentobarbital vs Thiopental & Hypotension during treatment \\ 
  Schwartz 1984 & Barbiturate vs Mannitol & Death at the end of follow-up (1 year) \\ 
  Schwartz 1984 & Barbiturate vs Mannitol & Death at the end of follow-up (1 year) \\ 
  Schwartz 1984 & Barbiturate vs Mannitol & Uncontrolled ICP during treatment \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Death at the end of follow-up \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Death or severe disability at the end of follow-up \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Mean ICP during treatment \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Mean arterial pressure during treatment \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Hypotension during treatment \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Mean body temperature during treatment \\ 
   \hline
\end{tabular}
\endgroup
\caption{Barbiturate and head injury review. In the columns, study names, comparison and outcome measure of the results are given} 
\label{barbiturates}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Thu Apr 25 14:26:47 2019
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lr}
  \hline
  \hline
Missing mean values and mean differences & 984 \\ 
  Missing standard deviations and standard errors & 1300 \\ 
  Missing sample sizes & 12173 \\ 
  Missing study year & 44649 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Number of missing variables and measurements in the dataset} 
\label{missing}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Thu Apr 25 14:26:47 2019
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lll}
  \hline
Outcome measure & n & Percentage \\ 
  \hline
Risk Ratio & 232583 & 50.1\% \\ 
  Mean Difference & 102315 & 22.1\% \\ 
  Odds Ratio & 49372 & 10.6\% \\ 
  Std. Mean Difference & 40535 & 8.7\% \\ 
  Peto Odds Ratio & 19122 & 4.1\% \\ 
  Hazard Ratio & 6566 & 1.4\% \\ 
  Risk Difference & 6234 & 1.3\% \\ 
  Rate Ratio & 2283 & 0.5\% \\ 
  other & 4810 & 1\% \\ 
   \hline
\end{tabular}
\endgroup
\caption{Frequencies of outcome measures among results. n denotes the total number 
             of results with the outcome measure and percentage the percentage of the outcome measure,} 
\label{outcome.measure.frequencies}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Thu Apr 25 14:26:47 2019
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lll}
  \hline
n & Number of groups & Cumulative sum of groups \\ 
  \hline
1 & 102344 & 186300 \\ 
  2 & 31686 & 83956 \\ 
  3 & 16072 & 52270 \\ 
  4 & 9628 & 36198 \\ 
  5 & 6444 & 26570 \\ 
  6 & 4230 & 20126 \\ 
  7 & 2961 & 15896 \\ 
  8 & 2114 & 12935 \\ 
  9 & 1592 & 10821 \\ 
  10 & 1238 & 9229 \\ 
  11 & 921 & 7991 \\ 
  12 & 702 & 7070 \\ 
  13 & 585 & 6368 \\ 
  14 & 455 & 5783 \\ 
  15 & 5328 & 5328 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Cumulative number of groups with number of reproduction trials >= n} 
\label{repr.groups}
\end{table}

@@

\section{Meta-analysis}
The data at hand is composed of over 400,000 comparisons with balanced sample size $n > 24$, which has been chosen as a threshold to assess the significance of treatment effect. The significance of the p-values of treatment effect estimates was calculated for all comparisons in the dataset with acceptable sample size and outcome measures as odds ratios, risk ratios, risk differences and mean and standardized mean differences. This led to somewhat more than 400,000 significance test results. Since most of the times, the study publication year is available, the fraction of significant treatment effects found is shown over time in Figure \ref{study.significance.overtime}.The $p$-value was chosen to be 0.05  only times where a reasonbly large number of effect estimates is available is shown in order to reduce random fluctuation ($n$ > 800).

@@

\end{knitrout}
\caption{Mean of the absolute value of the normalized effect size plotted against the total sample size.}
\label{study.significance.overtime}
\end{figure}

It is possible to analyse all studies with one or more replica by meta-analysis. There are different meta-analys methods that can be applied. While fixed effects meta-analysis pools the estimates and their variance, random effects meta-analysis additionally estimates the between study variances and adds this variance to the pooled within-study variances when assessing the uncertainity of the pooled effect estimate. The Hartung and Knapp adjustment of p-values and confidence intervals for random effects meta-analysis is more conservative than random-effects meta-analysis and will yieild less significant results than the latter.

\vspace{0mm}
In the following, the study results are combined with the results of their corresponding meta-analysis. Thus, every study that has one or more replicate was analysed by meta-analysis. In the following, the significance of the study, the primary significance, is compared with the significance of the meta-analysis or secondary significance.

\vspace{0mm}
In Figure \ref{primary.secondary.significance}, the overall fraction of studies with significant and non-significant meta-analysis results is shown. The meta-analysis method used to pool the treatment effect estimates and assessing significance is indicated on the right hand side of the Figure. One can see the overall fraction of rejected null-hypotheses of no treatment effect.

@@

One can visualise the fraction of studies with significant treatment effect separately for studies with significant primary treatment effect, i.e. non-pooled effect, and non-significant primary treatment effect. This can be seen in Figure \ref{primary.secondary.significance.sep.sig}. One can see that primary significance and non-significance is often overruled when performing a meta-analysis.

{\centering \includegraphics[width=\textwidth-3cm]{figure/ch02_figunnamed-chunk-15-1}

@@

The separation of studies can also be made based on the significance of heterogeneity between them when pooling them by means of a meta-analysis. Significant heterogeneity between studies corresponds to a rejection of the null-hypothesis that all the study treatment effect estimates share the same underlying distribution. The test used to assess heterogeneity was based on the between-study heterogeneity estimate $Q$ estimated as \citet{tau.estimator} suggested. This is shown in Figure \ref{primary.secondary.significance.sep.sig}.

@@

\end{knitrout}
\caption{Overall fraction of studies whose treatment effect estimate was significant when pooled by means of meta-analysis, separated by significance of study treatment effect estimate. The fractions have been calculated
by fixed-effects, random-effects and Hartung and Knapp adjusted random-effects meta-analysis}
\label{primary.secondary.significance.sep.sig}
\end{figure}

\section{Small study effects}
One crucial assumption in meta analysis is that the availability and publication of studies does not depend on their effect and the variance of the effect. If this is not given, one often speaks of publication bias. In fact, there can also be other reasons for this (see discussion section). A more appropriate term for the phenomenon is small study effect. If small study effects are present in a meta-analysis, the classical approaches to merge single study results in to an overall intervention effect fails to provide an appropriate estimate of the treatment effect.

\vspace{0mm}
To provide an overview over the issue, first it is shown how median absolute effect size decreases with increasing sample size of the comparisons in Figure \ref{effect.samplesize}. In some sense, this is the same idea as for a funnel plot, by depicting the size of effects relative to their variance.

\vspace{0mm}
A clear trend of decrease of absolute effect size with increasing sample size (i.e. smaller variance) is visible. It is particularly substantive from very small trials ($n$ = 10) to medium sample size ($n$ = 100) and afterwards it evens off. With increasing sample size, there are fewer results, therefore, the variation between medians increases. All effects are normalized by subtracting the mean effect size of the dataset and dividing through the standard deviation. Note that various types of outcome measures are included, such as mean difference and risk ratios, and are normalized with respect to all effects.

@@

The median absolute normalized effect size can be visualized for the different outcome measures separately. Then it is well visible in Figure \ref{effect.samplesize.separated} that the trend of effect size decrease holds in particualar for risk and odds ratios, while it is way more stable for mean differences and standardized mean differences. Instead of normalizing the effects (i.e. subtracting mean and dividing through the standard deviation) for all effect sizes, the effects are normalized with respect to the effects of the same outcome measures.

{\centering \includegraphics[width=\textwidth-3cm]{figure/ch02_figunnamed-chunk-18-1}

@@

% It is also possible to look how the amount of meta-analysis with significant pooled effect size estimates changes over time, e.g. mean publication year (of studies included in the meta-analysis). 0.05 is used here as the threshold for $p$-value significance and subsequent null-hypothesis rejection. To constrict random fluctuation in the proportion, only mean publication years with a certain amount of meta-analyses are shown ($n < 180$). The $p$-value of the fixed effects meta analysis model is used to decide about rejection of the null-hypothesis. As Figure \ref{treatment.effect.significance.overtime} shows, the proportion of meta-analyses with significant treatment effect estimates stays roughly constant over time. 
% 
% \begin{figure}
% <<echo=FALSE>>=
% meta %>% filter(mean.publication.year < 2013 & mean.publication.year > 1990) %>% 
%   ggplot(aes(x = mean.publication.year, fill = factor(sig.fixef), stat(count))) + geom_density(na.rm = T, position = "fill") + 
%   labs(fill = "Significance") + scale_fill_discrete(labels= c("Yes", "No"))
% @
% \caption{Proportion of significant treatment effect estimates over time (mean publication year).}
% \label{effect.samplesize}
% \end{figure}
%

@@

\vspace{0mm}
There are modifications to make tests more appropriate in case of binary outcomes, therefore the results have been separated in continuous and dichotomous outcome test results. In Figure \ref{bias.results.cont} the proportion of test results that led to rejection of the null hypothesis of no small study effect based on the 5 \% level are shown for continuous outcomes ($n$ = 1383)
The same is shown in Figure \ref{bias.results.bin} for dichotomous outcome measures ($n$ = 3442).

@@

@@

The same plots can be shown separately for meta-analyses with significant and non-significant pooled effect sizes. This is done in Figure \ref{bias.results.cont.sep} for continuous outcomes and \ref{bias.results.bin.sep} for binary outcomes.

{\centering \includegraphics[width=\textwidth-3cm]{figure/ch02_figunnamed-chunk-22-1}

\end{knitrout}
\caption{Proportion where the null hypothesis of no small study effect is rejected based on the 5\% significance level for different tests (continuous outcomes).}
\label{bias.results.cont.sep}
\end{figure}

{\centering \includegraphics[width=\textwidth-3cm]{figure/ch02_figunnamed-chunk-23-1}

\end{knitrout}
\caption{Proportion where the null hypothesis of no small study effect is rejected based on the 5\% significance level for different tests (dichotomous outcomes).}
\label{bias.results.bin.sep}
\end{figure}

Furthermore one can look if the frequencies of tests that reject the null hypotheses change over time (mean publication year of the studies included in the meta analyses). The proportion of the test results are shown in Figure \ref{pub.bias.time.overtime}. The Figure suggests that the frequency of publication bias remains constant over time. The mean publication years have been restricted such that at least 180 meta-analyses are available per year such that random fluctuation is restricted to some extent. The significance threshold for the $p$-values used is 0.05, and the small study effect test used is Thomson's test (with the arcsine variance stabilizing transformation function used in the case of binary outcomes).

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Thu Apr 25 14:27:29 2019
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lll}
  \hline
 & Test Agreement & P-value Correlation \\ 
  \hline
egger.schwarzer & 0.87 & 0.39 \\ 
  egger.peter & 0.89 & 0.51 \\ 
  egger.rucker & 0.88 & 0.48 \\ 
  egger.harbord & 0.93 & 0.76 \\ 
  schwarzer.peter & 0.89 & 0.29 \\ 
  schwarzer.rucker & 0.88 & 0.34 \\ 
  schwarzer.harbord & 0.92 & 0.51 \\ 
  rucker.peter & 0.92 & 0.67 \\ 
  harbord.peter & 0.91 & 0.52 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Proportion of tests that agree in rejection ar acceptance of the null hypothesis that there is no small study effect (dichotomous outcomes)} 
\label{agreement.bin}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Thu Apr 25 14:27:29 2019
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lll}
  \hline
 & Test Agreement & P-value Correlation \\ 
  \hline
thomson.egger & 0.89 & 0.69 \\ 
  thomson.begg & 0.82 & 0.47 \\ 
  egger.begg & 0.79 & 0.36 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Proportion of tests that agree in rejection ar acceptance of the null hypothesis that there is no small study effect (continuous outcomes)} 
\label{agreement.cont}
\end{table}

@@

@@

{\centering \includegraphics[width=\textwidth-3cm]{figure/ch02_figunnamed-chunk-33-1}

@@

{\centering \includegraphics[width=\textwidth-3cm]{figure/ch02_figunnamed-chunk-34-1}

@@

\section{Small Study Effects Tests}
\subsection{Continuous Outcome Tests}
% Essentially, there are two kinds of tests for reporting bias, non-parametrical or rank-based tests or regression based tests. 
% First, tests for continuous outcome studies are described, and special modifications of those for binary outcomes will be introduced later.

>>>>

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.98, 0.98, 0.98}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in UseMethod("{}mutate\_"{}): no applicable method for 'mutate\_' applied to an object of class "{}function"{}}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in pb.meta.merge(meta.bin, meta.cont): could not find function "{}pb.meta.merge"{}}}\end{kframe}
\end{knitrout}

@@

\section{Cochrane Systematic Reviews}
As has been mentioned before, the Cochrane Group has specialized on systematic reviews in clinical science. Certain knowledge of standards and principles of the Cochrane Group may help to assess the quality and the properties of the dataset. The following information stems from the Cochrane Handbook for Systematic Reviews \citep{cochrane.handbook}.

\vspace{0mm}
The definition of a systematic review is that it ``attempts to collate all empirical evidence that fits pre-specified eligibility criteria in order to answer a specific research question.'' Thus, the ``key properties of a review are'':

\begin{itemize}
\item``a clearly stated set of objectives with pre-defined eligibility criteria for studies''
\item ``an explicit, reproducible methodology''
\item ``a systematic search that attempts to identify all studies that would meet the eligibility criteria''
\item ``an assessment of the validity of the findings of the included studies, for example through the assessment of risk of bias''
\end{itemize}

At the end of a systematic review, ``a systematic presentation, and synthesis, of the characteristics and findings of the included studies'' is done.

\vspace{0mm}
53 Cochrane Review Groups prepare and maintain the reviews within specific areas of health care. A group consists of ``researchers, healthcare professionals and people using healthcare services (consumers)''.

\vspace{0mm}
The groups are supported by Method Groups, Centres and Fields. The Cochrane Method Groups aim to discuss and consult the groups in methodological questions concerning review preparation. The Centres play a main role in training and support of the Groups. The Fields are responsible for broad medical research areas and follow priorities in those areas by advice and control of the groups.

\vspace{0mm}
The first step in a review is writing a protocoll, specifying the research question, the methods to be used in literature search and analysis and the eligibility criteria of the study. Changes in protocolls are possible but have to be documented and the protocol is published in advance of the publication of the full review. The choices of methodology as well as the changes should not be made ``on the basis of how they affec the outcome of the research study''.

\vspace{0mm}
In order to avoid potential conflicts of interests, there is a code of conduct that all entities of the Cochrane Organization have to agree on: conflicts of interest must be disclosed and possibly be forwarded to the Cochrane Centre, and participation of review authors in the studies used have to be acknowledged. Additionally, a Steering Group publishes a report of potential conflicts of interests based on information about external funding of Cochrane Groups.

\vspace{0mm}
In order for keeping the reviews up-to-date, they are revised in a two-year cicle with exceptions. In addition to inclusion of new evidence in a field, the revision and maintenance process may as well includes change in analysis methods. This can reflect some advance in clinical science as for example new informations about important subgroups, as well as new methods for conducting a Cochrane Review. However, there are no clear guidelines and the Cochrane Groups are free in the rate and extent of up-dating their reviews.

\subsection{Methods for Cochrane Reviews}
A research question defines the following points: ``the types of population (participants), types of interventions (and comparisons), and the types of outcomes that are of interest''. From the research question, ususally the eligibility criteria follow. Usually, outcomes are not part of eligibility criteria, except for special cases such as adverse effect reviews.

\vspace{0mm}
The type of study is an important eligibility criterium. The Cochrane Collaboration focuses ``primarily on randomized controlled trials'', and also, the methods of study identification in literature search are focused on randomized trials. Furthermore, study characteristics such as blinding of study operators with respect to treatment and cluster-randomizing might be additional eligibility criteria which have to be chosen by the review authors.

\vspace{0mm}
After having specified the eligibility criteria, studies have to be collected. The central idea of systematic reviews, and also meta-analyses, is that the collected studies are a random sample of a population of studies, i.e. that they are representative and can be used to assess population properties. Therefore, the search process is crucial, as a selective search result may impose bias on the sample of studies available, making it a non-random sample. For this purpose, the Cochrane Groups are advised to go beyond MEDLINE !!cite!!, because a search restricted to it has been shown to deliver only 30\% to 80\% of available studies. ``Time and budget restraints require the review author to balance the thoroughness of the search with efficiency in use of time and funds and the best way of achieving this balance is to be aware of, and try to minimize, the biases such as publication bias and language bias that can result from restricting searches in different ways.'' It is important to note that not only studies, but also study reports are occasionally used in the reviews, as they may provide useful information.

\vspace{0mm}
There are different sources that are being used to search for studies. 
\begin{itemize}
\item The Cochrane Central Register of Controlled Trials is a source of reports of controlled trials. ``As of January 2008 (Issue 1, 2008), CENTRAL contains nearly 530,000 citations to reports of trials and other studies potentially eligible for inclusion in Cochrane reviews, of which 310,000 trial reports are from MEDLINE, 50,000 additional trial reports are from EMBASE and the remaining 170,000 are from other sources such as other databases and handsearching.'' It includes citations published in many languages, citations only available in conference proceedings, citations from trials registers and trials results registers.
\item MEDLINE. MEDLINE includes over 16 million references to journal articles. 5,200 journals publishing in 27 languages are indexed for MEDLINE. PubMed gives access to a free version of MEDLINE with up-to-date citations. NLM gateway such as the Health Services Research Project, Meeting Abstracts and TOXLINE Subset for toxicology citations allows for search in both databases together with additional data from the US National Library of Medicine.
\item EMBASE. 4,800 Journals publishing in 30 languages are indexed to EMBASE, which includes more than 11 million records from 1974 onwards. EMBASE.com also includes 7 million unique records from MEDLINE (1966 up to date) together with its own records. Additionally, EMBASE Classic allows access to digitized records from 1947 to 1973. EMBASE and MEDLINE each have around 1,800 journals not indexed in the other databse.
\item Regional or national and subject specific databases can additionally be consulted and often provide important information. Financial considerations may limit the use of such databases.
\item General search engines such as Google Scholar, Intute and Turning Research into Practice (TRIP) database can be used.
\item Citation Indexes. The database lists articles published in around 6,000 Journals with articles in which they have been cited and is available online as SciSearch. This form of search is known as cited reference searching.
\item Dissertation sources. Dissertations are often listed in MEDLINE or EMBASE but one is advised to also search in specific dissertation sources.
\item Grey Literature Databases. Approximately 10\% of the results in the Cochrane Database stems from conference abstracts and other grey literature. The Institute for Scientific and Technical Information in France provides access to entries of the previously closed System for Information on Grey Literature database of the European Association for Grey Literature Exploitation). Another source is the Healthcare Management Information Consortium (HMIC) database containing records from the Library and Information Services department of the Department of Health (DH) in England and the King's Fund Information and Library Service. The National Technical Information Service (NTIS) gives access to the results of US and non-US government-sponsored research, as well as technical report for most published results. References from newsletters, magazines and technical and annual reports in behavioral science, psychology and health are provided in the PsycEXTRA database which is linked to PsycINFO database.
\end{itemize}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Apr 29 12:57:33 2019
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{lllrrrr}
  \hline
Study & Comparison & Outcome & Events & Total & Events\_c & Total\_c \\ 
  \hline
Bohn 1989 & Barbiturate vs no barbiturate & Death at the end of follow-up & 11 & 41 & 11 & 41 \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Death at the end of follow-up & 14 & 27 & 13 & 26 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Example of two results as given in the dataset. Events denotes the count of events in the treatment group while Events c the count of events in the group compared to. Further descriptive variables have been ommitted} 
\label{barbiturate.row}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Apr 29 12:57:33 2019
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lll}
  \hline
Study & Comparison & Outcome \\ 
  \hline
Bohn 1989 & Barbiturate vs no barbiturate & Death at the end of follow-up \\ 
  Bohn 1989 & Barbiturate vs no barbiturate & Death or severe disability at the end of follow-up \\ 
  Eisenberg 1988 & Barbiturate vs no barbiturate & Death at the end of follow-up \\ 
  Eisenberg 1988 & Barbiturate vs no barbiturate & Uncontrolled ICP during treatment \\ 
  Eisenberg 1988 & Barbiturate vs no barbiturate & Hypotension during treatment \\ 
  Perez-Barcena 2008 & Pentobarbital vs Thiopental & Death at the end of follow-up (6 months) \\ 
  Perez-Barcena 2008 & Pentobarbital vs Thiopental & Death or severe disability at the end of follow-up (6 months) \\ 
  Perez-Barcena 2008 & Pentobarbital vs Thiopental & Uncontrolled ICP during treatment \\ 
  Perez-Barcena 2008 & Pentobarbital vs Thiopental & Hypotension during treatment \\ 
  Schwartz 1984 & Barbiturate vs Mannitol & Death at the end of follow-up (1 year) \\ 
  Schwartz 1984 & Barbiturate vs Mannitol & Death at the end of follow-up (1 year) \\ 
  Schwartz 1984 & Barbiturate vs Mannitol & Uncontrolled ICP during treatment \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Death at the end of follow-up \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Death or severe disability at the end of follow-up \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Mean ICP during treatment \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Mean arterial pressure during treatment \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Hypotension during treatment \\ 
  Ward 1985 & Barbiturate vs no barbiturate & Mean body temperature during treatment \\ 
   \hline
\end{tabular}
\endgroup
\caption{Barbiturate and head injury review. In the columns, study names, comparison and outcome measure of the results are given} 
\label{barbiturates}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Apr 29 12:57:33 2019
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lr}
  \hline
  \hline
Missing mean values and mean differences & 984 \\ 
  Missing standard deviations and standard errors & 1300 \\ 
  Missing sample sizes & 12173 \\ 
  Missing study year & 44649 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Number of missing variables and measurements in the dataset} 
\label{missing}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Apr 29 12:57:33 2019
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lll}
  \hline
Outcome measure & n & Percentage \\ 
  \hline
Risk Ratio & 232583 & 50.1\% \\ 
  Mean Difference & 102315 & 22.1\% \\ 
  Odds Ratio & 49372 & 10.6\% \\ 
  Std. Mean Difference & 40535 & 8.7\% \\ 
  Peto Odds Ratio & 19122 & 4.1\% \\ 
  Hazard Ratio & 6566 & 1.4\% \\ 
  Risk Difference & 6234 & 1.3\% \\ 
  Rate Ratio & 2283 & 0.5\% \\ 
  other & 4810 & 1\% \\ 
   \hline
\end{tabular}
\endgroup
\caption{Frequencies of outcome measures among results. n denotes the total number 
             of results with the outcome measure and percentage the percentage of the outcome measure,} 
\label{outcome.measure.frequencies}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Apr 29 12:57:33 2019
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lll}
  \hline
n & Number of groups & Cumulative sum of groups \\ 
  \hline
1 & 102344 & 186300 \\ 
  2 & 31686 & 83956 \\ 
  3 & 16072 & 52270 \\ 
  4 & 9628 & 36198 \\ 
  5 & 6444 & 26570 \\ 
  6 & 4230 & 20126 \\ 
  7 & 2961 & 15896 \\ 
  8 & 2114 & 12935 \\ 
  9 & 1592 & 10821 \\ 
  10 & 1238 & 9229 \\ 
  11 & 921 & 7991 \\ 
  12 & 702 & 7070 \\ 
  13 & 585 & 6368 \\ 
  14 & 455 & 5783 \\ 
  15 & 5328 & 5328 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Cumulative number of groups with number of reproduction trials >= n} 
\label{repr.groups}
\end{table}

@@

\section{Meta-analysis}
The data at hand is composed of over 400,000 results with balanced sample size $n > 24$. The significance of the p-values of treatment effect estimates was calculated for all results in the dataset with acceptable sample size and outcome measures as odds ratios, peto odds ratios, risk ratios, risk differences and mean and standardized mean differences.

%Significance over time? Chunck 1

It is possible to analyse all results that are not unique in their comparison, outcome and subgroup in the dataset by meta-analysis. So each study that has been repeated with respect to experimental setup is included. There are different meta-analys methods that can be applied, here, fixed-effects meta-analysis, random-effects meta-analysis and random-effects meta-analysis using the Hartung-Knapp method are used.

\vspace{0mm}
In the following, the results are combined with the results of their corresponding meta-analysis if possible (i.e. if studies were not unique in the dataset with respect to their experimental setup). The significant of the treatment effect in a single result is then compared to the significance of the pooled treatment effect. 
In Figure \ref{primary.secondary.significance}, the overall fraction of studies with significant and non-significant meta-analysis results is shown. The meta-analysis method used to pool the treatment effect estimates and assessing significance is indicated on the right hand side of the Figure. The colors indicate the overall fraction of rejected or accepted null-hypotheses of no treatment effect.

@@

One can visualise the fraction of studies with significant treatment effect separately for studies with significant primary treatment effect, and non-significant primary treatment effect, as done in Figure \ref{primary.secondary.significance.sep.sig}. It can be seen that primary significance and non-significance is often overruled when performing a meta-analysis.

@@

@@

The amount of significant heterogeneity in the results used by the meta-analyses is 53350 and corresponds to 81.5936377\% of the total number of meta-analyses.

%Comparison of significance separated for significant Q? Chunck 2

%Proportion of significant pooled estimates based on proportion of single pooled estimates? Chunck 5

\section{Small study effects}

To provide an overview over the abundance of small study effects in the dataset, first it is shown how median absolute effect size decreases with increasing sample size of the results (Figure \ref{effect.samplesize}). 
%In some sense, this is the same idea as for a funnel plot, by depicting the size of effects relative to their variance.

\vspace{0mm}
A clear trend of decrease of absolute effect size with increasing sample size (i.e. smaller variance) is visible. 
%It is particularly substantive from very small trials ($n$ = 10) to medium sample size ($n$ = 100) and afterwards it evens off. With increasing sample size, there are fewer results, therefore, the variation between medians increases. 
All effects are normalized by subtracting the mean effect size of the dataset and dividing through the standard deviation. Note that various types of outcome measures are included, such as mean difference and risk ratios, and are normalized with respect to all effects.

@@

The median absolute normalized effect size can be visualized for the different outcome measures separately (Figure \ref{effect.samplesize.separated}). The plot confirms the trend of median effect size decrease towards lower sample size (e.g. for Risk Ratios: decrease $> 0.1$ standard deviations from 10 to 50 study participants). Instead of normalizing the effects 
%(i.e. subtracting mean and dividing through the standard deviation) 
for all effect sizes, the effects are normalized with respect to the effects of the same outcome measures.

@@

@@

%Show absolute effect size depending on scaled time? Chunck 3

@@

\vspace{0mm}
There are modifications to make tests more appropriate in case of binary outcomes, therefore the results have been separated in continuous and dichotomous outcome test results. In Figure \ref{bias.results.cont} the proportion of test results that led to rejection of the null hypothesis of no small study effect based on the 5 \% level are shown for continuous outcomes ($n$ = 1383)
The same is shown in Figure \ref{bias.results.bin} for dichotomous outcome measures ($n$ = 3442). The amount of studies varies from 5\% (Schwarzer's Test) to 13 \% (Egger's Test) for binary outcomes and 9\% (Begg and Mazumdar's Test) to 25 \% (Egger's Test) for continuous outcomes.

{\centering \includegraphics[width=\textwidth-3cm]{figure/ch02_figunnamed-chunk-19-1}

@@

There is no substantive change in the fraction of positive test results, depending on if the pooled treatment effect size estimate is significant or not. Most substantively, the fraction of positive Egger Test results increases for significant pooled treatment effects by 0.0768414 (the minimal increase is 0.0103761 for Schwarzer's Test).

%Show plots of significant meta-analyses separated? Chunck 4

Furthermore one can look if the frequencies of tests that reject the null hypotheses change over time (mean publication year of the studies included in the meta analyses). The proportion of the test results are shown in Figure \ref{pub.bias.time.overtime}. The Figure suggests that the frequency of publication bias remains constant over time. The mean publication years have been restricted such that at least 180 meta-analyses are available per year, such that random fluctuation is restricted to some extent. The significance threshold for the $p$-values used is 0.05, and the small study effect test used is Thomson's test (with the arcsine variance stabilizing transformation function used in the case of binary outcomes).

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Apr 29 12:58:07 2019
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lll}
  \hline
 & Test Agreement & P-value Correlation \\ 
  \hline
egger.schwarzer & 0.87 & 0.39 \\ 
  egger.peter & 0.89 & 0.51 \\ 
  egger.rucker & 0.88 & 0.48 \\ 
  egger.harbord & 0.93 & 0.76 \\ 
  schwarzer.peter & 0.89 & 0.29 \\ 
  schwarzer.rucker & 0.88 & 0.34 \\ 
  schwarzer.harbord & 0.92 & 0.51 \\ 
  rucker.peter & 0.92 & 0.67 \\ 
  harbord.peter & 0.91 & 0.52 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Proportion of tests that agree in rejection ar acceptance of the null hypothesis that there is no small study effect (dichotomous outcomes)} 
\label{agreement.bin}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Apr 29 12:58:07 2019
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lll}
  \hline
 & Test Agreement & P-value Correlation \\ 
  \hline
thomson.egger & 0.89 & 0.69 \\ 
  thomson.begg & 0.82 & 0.47 \\ 
  egger.begg & 0.79 & 0.36 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Proportion of tests that agree in rejection ar acceptance of the null hypothesis that there is no small study effect (continuous outcomes)} 
\label{agreement.cont}
\end{table}

@@

@@

{\centering \includegraphics[width=\textwidth-3cm]{figure/ch02_figunnamed-chunk-25-1}

@@

Because the real amount of publication bias in the dataset is not known, the correction method can also be applied only to meta-analyses that have publication bias according to the small study effect tests in the previous section. Because the test developed by \citet{thomson.sharp} has been applied to both binary and continuous outcome meta-analyses (in the case of binary outcomes to arcsine variance stabilized proportions), it is used as a criterium to distinguish biased from unbiased meta-analyses. The proportions of significance tests of pooled treatment effects that turned from significant to non-significant, non-significant to significant etc. are shown in Figure \ref{significance.change.fixed.sep}. Fixed effects meta-analysis has been used to determine signficance of the uncorrected estimate.

@@

@@

\section{Small Study Effects Tests}
One crucial assumption in meta analysis is that the availability and publication of studies does not depend on their effect and the variance of the effect. If this is not given, one often speaks of publication bias. In fact, there can also be other reasons for this (see discussion section). A more appropriate term for the phenomenon is small study effect. If small study effects are present in a meta-analysis, the classical approaches to merge single study results in to an overall intervention effect fails to provide an appropriate estimate of the treatment effect.
